### 2015-07-05 presentation/presentation.tex
Add basic skeleton for the presentation

\documentclass[12pt, utf8, ngerman]{beamer}

\include{rvs}

\title{Analyse von Common Crawl}
\author{Erik Körner \and Kai Hainke \and Klemens Schölhorn}
\date{06.07.2015}

\begin{document}

% Titlepage
{
    \setbeamertemplate{footline}[default]
    \begin{frame}
        \titlepage
    \end{frame}
}

\begin{frame}{Übersicht}
    \tableofcontents
\end{frame}

\section{Common Crawl}

\subsection{Organisation}
\begin{frame}{\insertsection: \insertsubsection}
\end{frame}

\subsection{Technik}
\begin{frame}{\insertsection: \insertsubsection}
\end{frame}

\section{Analyse}

\subsection{Daten}
\begin{frame}{\insertsection: \insertsubsection}
\end{frame}

\subsection{Extraktion}
\begin{frame}{\insertsection: \insertsubsection}
\end{frame}

\section{Ergebnisse}

\subsection{TLD}
\begin{frame}{\insertsection: \insertsubsection}
\end{frame}

\subsection{Public Suffix}
\begin{frame}{\insertsection: \insertsubsection}
\end{frame}

\subsection{HTTPS}
\begin{frame}{\insertsection: \insertsubsection}
\end{frame}

\subsection{Webserver}
\begin{frame}{\insertsection: \insertsubsection}
\end{frame}

\section{Textextraktion}

\begin{frame}{\insertsection}
\end{frame}

\end{document}